---
title: "Overview of loopcityData"
output:
  BiocStyle::html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Overview of loopcityData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

`loopcityData` provides the experimental data used in examples and vignettes
for the [loopcity](https://sarmapar.com/loopcity/) package.

Two datasets ship directly with the package (accessible via `data()`). Larger
files — K562 Hi-C contact matrices and ChIP-seq bigWig tracks — are hosted on
ExperimentHub and downloaded to a local cache on first use.

## Shipped datasets

```{r shipped}
library(loopcityData)

data(GM12878_10KbLoops)
GM12878_10KbLoops

data(GM12878_10KbLoopsChr21)
GM12878_10KbLoopsChr21
```

## ExperimentHub datasets

The K562 datasets from Bond et al. 2023 are retrieved via ExperimentHub.
Files are downloaded once and cached locally by BiocFileCache.

```{r hub, eval=FALSE}
hic_path     <- K562_hic()
loops_path   <- K562_loops()
ctcf_path    <- K562_CTCF_bw()
h3k27ac_path <- K562_H3K27ac_bw()
```

These paths can be passed directly to `loopcity` functions and
`plotgardener` visualization functions. See the
[plotgardener integration vignette](https://sarmapar.com/loopcity/articles/
plotgardener-integration.html) in the `loopcity` package for a worked example.

## Data provenance

| Object | Accession | Reference |
|---|---|---|
| `GM12878_10KbLoops` / `GM12878_10KbLoopsChr21` | GSE63525 | Rao et al. 2014 |
| `K562_hic` / `K562_loops` | GSE214123 | Bond et al. 2023 |
| `K562_CTCF_bw` / `K562_H3K27ac_bw` | GSE213908 | Bond et al. 2023 |

## Session info

```{r session}
sessionInfo()
```
